---
title: false
output: html_document
params:
  param_namesOE: "default"
  param_namesHmgzd: "default"
  param_rowCount: "default"
  param_locationVars: "default"
  param_numProfileOE: "default"
  param_numProfileHZ: "default"
  param_profileVars: "default"
  param_summaryPlots: "default"
---

#### data homogenization processing summary and QC check results

**files processed: **
```{r filesProcessed, echo=FALSE, results='asis'}

set.alignment('left')
bind_rows(
  tibble(
    type = "provided data",
    filename = params$param_namesOE
  ),
  tibble(
    type = "homogenized data",
    filename = params$param_namesHmgzd
  )) %>% 
pandoc.table(split.table = Inf, missing = "")
```


**summary of row counts from submitted and homogenized data: **
```{r rowCount, echo=FALSE, results='asis'}

set.alignment('left')
pandoc.table(params$param_rowCount, split.table = Inf, missing = "")
# pandoc.table(data.frame(params$param_rowCount), split.table = Inf, missing = "")
```


**confirm all location-level variables entered into the key file were included in the homogenized data**
```{r locationDataCheck, echo=FALSE, results='asis'}

set.alignment('left')
pandoc.table(params$param_locationVars, split.table = Inf, missing = "")
```


**confirm all profile-level variables entered into the key file were included in the homogenized data**

*profile data variables expected:* `r params$param_numProfileOE`  
*profile data variables homogenized:* `r params$param_numProfileHZ`

```{r profileDataSummary, echo=FALSE, results='asis'}

set.alignment('left')
pandoc.table(params$param_profileVars, split.table = Inf, missing = "")
```


#### summary plotting

```{r summaryPlotting, echo=FALSE, warning=FALSE}

print(params$param_summaryPlots)
```
